Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-05-29 manuscript.Rmd
Edit: Present study section

## The Present Work
Social media has become a primary platform for social movements and
activism across the world, but the extent to which virtual activism
reflects real-world mobilization is still largely unknown. We
conceptualize the physical and virtual movements as a dynamical system,
both constantly interacting and influencing the other. Both real- and
virtual-world collective coordination were powerful tools in the events
of Arab Spring, and are undeniably linked. This study aims to explore
the dynamics of Syrian individuals who took to Twitter to raise
awareness about the events of Arab Spring and bring about mass mobilization.
Using cross-recurrence quantification analysis [CRQA; @zbilut1998detecting]
and windowed cross-recurrence quantification analysis
[WCRQA; @webber2005recurrence], we analyzed the dynamics of social
cohesion of tweets from March 30, 2012 to June 15, 2012 and the counts
of all, positive, and negative real-world events on those respective days.
We hypothesized that the count of all events and the count of negative
events would show high levels of determinism, as this time period was
marked by daily violence which often leads to increased salience. We also
hypothesized that the count of positive events would not show patterns of
determinism, as these events were likely overshadowed by the salient
negative daily occurrences. Using WCRQA, we hoped to identify shifts in the
dynamics of the systems that could then be linked to real world shifts in
conflict frequency. As social cohesion has never been analyzed using
nonlinear methods, the analyses contained herein are largely exploratory.
The goal of this endeavor is to gain a better understanding of how events
of different intensities during times of turmoil may enhance or deter
social cohesion in social media communication, and how social mobilization
may be evident in the dynamic coupling of cohesion and real-world events.

>>>>

## The Present Work
Social media has become a primary platform for social movements and
activism across the world, but the extent to which virtual activism
reflects and drives real-world mobilization is still largely unknown. 
In the current work, we
conceptualize the physical and virtual movements as a dynamical system,
both constantly interacting and influencing the other. We aim to study
this by analyzing the real-world events and online social cohesion
during nearly three months of the Arab Spring. Both real- and
virtual-world collective coordination were powerful tools in the events
of the Arab Spring and are undeniably linked. This study aims to explore
the dynamics of Syrian Twitter users, who used the social media platform to raise
awareness about the Arab Spring and bring about mass mobilization.

Using cross-recurrence quantification analysis [CRQA; @zbilut1998detecting]
and windowed cross-recurrence quantification analysis
[WCRQA; @webber2005recurrence], we analyzed the co-evolution of social
cohesion online and real-world events. Specifically, using heterogeneous
data from March 30, 2012, to June 15, 2012, we compared the social cohesion
of Syrian tweets with the counts
of all, positive, and negative real-world events.

We hypothesized that the count of all events and the count of negative
events would show high levels of determinism, as this time period was
marked by daily violence which often leads to increased salience. We also
hypothesized that the count of positive events would not show patterns of
determinism, as these events were likely overshadowed by the salient
negative daily occurrences. 
After testing these hypotheses, we then conducted exploratory analyses 
to further examine the patterns identified with CRQA. Using WCRQA,
we sought to identify shifts in the
dynamics of the systems that could then be linked to real-world shifts in
conflict frequency.
Through these analyses, our primary goal in the current work was to uncover
(1) how different intensities during times of turmoil may enhance or deter
social cohesion in social media communication and (2) how social mobilization
may be evident in the dynamic coupling of cohesion and real-world events.

An additional focus of the current work is to introduce cross-recurrence
quantification analysis---a longstanding method from physics that has 
since become influential in a variety of other fields---to political 
science. We do so by providing detailed descriptions of the methods and
providing clear directions about interpretation and significance testing
even with relatively small-*n* or case studies. After demonstrating these
methods' utility with this dataset, we close the paper with specific
suggestions to political scientists interested in incorporating these
methods into their work.
